---
title: "CFA Script"
output: html_document
date: "2024-07-10"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Dr Paul Christiansen
CFA

Packages to install- only needs to be done once.

```{r}

#install.packages("lavaan")      ### does LAtent VAriable ANalysis see https://lavaan.ugent.be/
#install.packages("lavaanPlot")  ### make plots https://cran.r-project.org/web/packages/lavaanPlot/vignettes/Intro_to_lavaanPlot.html
library(lavaan)
library(lavaanPlot)
```

Turn off scientfic notation

```{r}
options(scipen = 999)
```

Read in the data set

```{r}

cfadata <- read.csv("C:/Users/kiera/Downloads/cfaimputedasfactor.csv")
View(cfadata)   
str(cfadata)
```






Model 1: EFA1 suggested
```{r}

###fit your model, the model need to be within 'single quotations' and will only run if text is green! 
###F1 is simply the name I have given the the factor you can call it whatever you like 
### The =~ (equals tilde) symbol means measured by so factor one is measured by ATGC1, ATGC2 and ATGC3 and so on

model1 <- 'NeutAnx=~NIES4+NIES9+NIES10+NIES11+NIES12+NIES16+NIES17+NIES18+NIES19
AnxSig=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
DL=~NIES3+NIES5+NIES6'

  
fit1 <- cfa(model1, estimator = "WLSMV", data = cfadata, std.lv = TRUE)    

fit1 <- cfa(model1, estimator = "WLSMV", data = cfadata, std.lv = TRUE,

ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6")) #### fit the above model             
summary(fit1, standardized=TRUE, ci=TRUE, fit.measures=TRUE) #### display the results. Standardised = TRUE gives beta values (std.all column), ci=TRUE gives 95%CIs
modindices(fit1, minimum.value = 10, sort = TRUE)

```

Model 2: EFA1 Adapted 
```{r}
model2 <- 'NeutAnx=~NIES9+NIES10+NIES11+NIES12+NIES16+NIES17+NIES18+NIES19
AnxSig=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
DL=~NIES4+NIES3+NIES5+NIES6'

  
fit2 <- cfa(model2, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))                                 
summary(fit2, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit2, minimum.value = 10, sort = TRUE)
```

Model 3: EFA2 suggested 
```{r}
model3 <- 'Anx=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
Gastric=~NIES4+NIES9+NIES10+NIES16+NIES17
Thirst=~NIES3+NIES5+NIES6+NIES11+NIES12+NIES18+NIES19'

  
fit3 <- cfa(model3, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))                                 
summary(fit3, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit3, minimum.value = 10, sort = TRUE)
```

Model 4: EFA2 Adapted
```{r}
model4 <- 'Anx=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
Gastric=~NIES3+NIES4+NIES9+NIES10+NIES16+NIES17
Thirst=~NIES5+NIES6+NIES11+NIES12+NIES18+NIES19'

  
fit4 <- cfa(model4, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))
summary(fit4, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit4, minimum.value = 10, sort = TRUE)
```
Model 5: EFA2 Adapted correlated residuals added 
```{r}
model5 <- 'Anx=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
Gastric=~NIES3+NIES4+NIES9+NIES10+NIES16+NIES17
Other=~NIES5+NIES6+NIES11+NIES12+NIES18+NIES19
NIES7 ~~ NIES8'

  
fit5 <- cfa(model5, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))
summary(fit5, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit5, minimum.value = 10, sort = TRUE)
```


Model 6: EFA2 Adapted pain removed
```{r}
model6 <- 'Anx=~NIES1+NIES2+NIES7+NIES8+NIES14+NIES15
Gastric=~NIES3+NIES4+NIES9+NIES10+NIES16+NIES17
Thirst=~NIES5+NIES6+NIES11+NIES12+NIES18+NIES19'

  
fit6 <- cfa(model6, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))
summary(fit6, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit6, minimum.value = 10, sort = TRUE)

```

Model 7: EFA2 Adapted and don't like removed
```{r}
model7 <- 'Anx=~NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
Gastric=~NIES9+NIES10+NIES16+NIES17
Thirst=~NIES11+NIES12+NIES18+NIES19'

  
fit7 <- cfa(model7, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))                                 
summary(fit7, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit7, minimum.value = 10, sort = TRUE)
```

Model 8: 2 FACTOR
```{r}
model8 <- 'Anx=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20
Neut=~NIES3+NIES4+NIES5+NIES6+NIES9+NIES10+NIES11+NIES12+NIES16+NIES17+NIES18+NIES19'

  
fit8 <- cfa(model8, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))                                 
summary(fit8, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit8, minimum.value = 10, sort = TRUE)
```
Model 9: 1-factor model
```{r}
model9 <- 'Fact=~NIES1+NIES2+NIES7+NIES8+NIES13+NIES14+NIES15+NIES20+NIES3+NIES4+NIES5+NIES6+NIES9+NIES10+NIES11+NIES12+NIES16+NIES17+NIES18+NIES19'

  
fit9 <- cfa(model9, estimator = "WLSMV", data = cfadata, std.lv = TRUE, ordered = c("NIES4","NIES9","NIES10","NIES11","NIES12","NIES16",

"NIES17","NIES18","NIES19","NIES1","NIES2","NIES7",

"NIES8","NIES13","NIES14","NIES15","NIES20","NIES3","NIES5","NIES6"))                                 
summary(fit9, standardized=TRUE, ci=TRUE, fit.measures=TRUE)
modindices(fit9, minimum.value = 10, sort = TRUE)
```


########Note on fitting methods

#### this fits a maximum likelihood model by default  in code it looks like
###fit <- cfa(model,estimator = "ML", data = ATGC)

## Lavaan alows lots of different estimators
# "GLS": generalized least squares. For complete data only.
#"WLS": weighted least squares (sometimes called ADF estimation). For complete data only.
#"DWLS": diagonally weighted least squares
#"ULS": unweighted least squares

#Many estimators have 'robust' variants, meaning that they provide robust standard errors and a scaled test:
#"MLM": maximum likelihood estimation with robust standard errors and a Satorra-Bentler scaled test
#statistic. For complete data only.
#"MLMVS": maximum likelihood estimation with robust standard errors and a mean- and variance adjusted
#test statistic (aka the Satterthwaite approach). For complete data only.
#"MLMV": maximum likelihood estimation with robust standard errors and a mean- and variance adjusted
#test statistic (using a scale-shifted approach). For complete data only.
#"MLF": for maximum likelihood estimation with standard errors based on the first-order derivatives, and
#conventional test statistic. For both complete and incomplete data.
#"MLR": maximum likelihood estimation with robust (Huber-White) standard errors and a scaled test
#statistic that is (asymptotically) equal to the Yuan-Bentler test statistic. For both complete and incomplete data

#modification indicies tell you if your model can be improved by using covariances, 
#a modification index is the chi squared reduction adding a covariance will have, 
# we have set it to ten. Covariance should only be added were logical  


##################################plot

#################
####this package allows us creae diagrams
### all of these things are standard, you fit the model you have created 'fit2' in this case. I have asked for coefficients
### coefs=T and for these to be standardised stand =2.
### if you want covariances between facts add ,covs=T, 

```{r}


lavaanPlot(model = fit4, node_options = list(shape = "box", fontname = "Helvetica"), edge_options = list(color = "black"),
           coefs = T, stand=T, covs = T)
```